### 2023-09-19 assignment_03stat405.qmd
Add files via upload

---
title: 'Assignment 3 STAT 405'
format:
  pdf: default
pdf-engine: pdflatex  # You may need to remove this line
---

* Please produce your assignment as a pdf (knit to pdf).
See instructions in announcements if you have not downloaded a LaTeX distributor.

* Please also submit your
qmd file (it will not be graded but we want it for reference
purposes).
* Please use the provided qmd file as a basis for your first assignment
(do not submit an R script file), putting your solutions
inside the empty blocks that follow "Solution:"
* You will have to produce your future qmd files.
* 20 points per exercise

<br />

#### Exercise 1:
Using paste (or paste0), create the following Latex formula: a_{1}X_{1}ˆ0 + a_{2}X_{2}ˆ1+ a_{3}X_{3}ˆ2 + a_{4}X_{4}ˆ3 + a_{5}X_{5}ˆ4

Solution:
```{r}
paste0("a_{",1:5,"}X_{",1:5,"}^",0:4, collapse = "+")
```

<br />

#### Exercise 2:
A bucket contains 12 rocks: five yellow rocks and seven black rocks. If we select three rocks at random, what is the probability that exactly one yellow rock gets selected? 
```{r}
#1. (9 points) using a for loop with an if() clause and accumulator (Hint: The strategy is to sample 3 rocks without replacement from sample_space
sample_space <- c(rep("Yellow", 5), rep("Blue", 7))
n <- 10000
set.seed(123)
accumulator <- 0
for(i in 1:n){
  curr_sample <- sample(sample_space, 3)
  if( "Yellow" %in% curr_sample && table(curr_sample)[["Yellow"]] == 1 )
    accumulator <- accumulator +1
}
accumulator/n

#2 (8 points) using a for loop with an ifelse() function and accumulator
set.seed(123)
accumulator <- 0
for(i in 1:n){
  curr_sample <- sample(sample_space, 3)
  ifelse( "Yellow" %in% curr_sample && table(curr_sample)[["Yellow"]] == 1,  accumulator <- accumulator +1, next)
}
accumulator/n

#3. (8 points) Vectorial approach using sapply (no loops, no if() or ifelse(), no accumulator).
set.seed(123)
simulated_results <- sapply(1:n, function(x) {
  curr_sample <- sample(sample_space, 3)
  num_yellow_selected <- sum(curr_sample == "Yellow")
  return(num_yellow_selected)
})

simulated_probability <- sum(simulated_results == 1) / n
simulated_probability
```

<br />

#### Exercise 3:
1. (12.5 points) In the previous exercise, what is the probability that at most one yellow rock gets selected?
2. (12.5 points) In the previous exercise, what is the probability that at least one yellow rock gets selected?

Solution:
```{r}
#1
set.seed(123)
simulated_results <- sapply(1:n, function(x) {
  curr_sample <- sample(sample_space, 3)
  num_yellow_selected <- sum(curr_sample == "Yellow")
  return(num_yellow_selected)
})

simulated_probability <- sum(simulated_results == 1 | simulated_results == 0) / n
simulated_probability

#2

set.seed(123)
simulated_results <- sapply(1:n, function(x) {
  curr_sample <- sample(sample_space, 3)
  num_yellow_selected <- sum(curr_sample == "Yellow")
  return(num_yellow_selected)
})

simulated_probability <- sum(simulated_results == 1 | simulated_results == 2 | simulated_results == 3) / n
simulated_probability

```
#### Exercise 4:

1. (9 points) Suppose you take $1000 to the casino and you play roulette. Every turn, you decide to bet 100 dollars on black. If you win, which occurs with probability 9/19, you win 200, which is a net profit of 100. If you lose, you get nothing. Using a while loop, write code to simulate your adventure until you lose all your money. How many rounds will it take?
2. (8 points) Now, repeat the above simulation 1,000 times. What is the average number of times you get to play for each $1000 bankroll you use? Plot a histogram of the number of times you get to play for each bankroll. Comment a sentence on the histogram
3. (8 points) Suppose you quit if you double your money. Estimate the probability you double your money by running the simulation 1,000 times.
```{r}
#1
set.seed(123)
play_roulette <- function(x) {
money <- 1000
rounds <- 0
outcomes <- c("win", "lose")
  while(money >0){
  rounds<- rounds + 1
  money <- money - 100
  play_outcome<- sample(outcomes, size = 1, prob=c(9/19, 10/19))
  if(play_outcome == "win"){
    money <- money + 200
    }
  }
return(rounds)
}

print(play_roulette())
#2
set.seed(123)
num_sim <- 1000
roulette_sim <- sapply(1:num_sim, play_roulette)
mean(roulette_sim)
hist(roulette_sim, main = "Histogram of Roulette Simulation", xlab = "Rounds")

#This histogram shows that most rounds will be in the lower range of possible 
#rounds/is right skewed.Very unlikely to play a lot of rounds.

#3
set.seed(123)
play_double_or_nothing_roulette <- function(x) {
money <- 1000
rounds <- 0
outcomes <- c("win", "lose")
  while(money >0 && money < 2000){
    rounds<- rounds + 1
    money <- money - 100
    play_outcome<- sample(outcomes, size = 1, prob=c(9/19, 10/19))
    if(play_outcome == "win"){
      money <- money + 200
    }
  }
  if(money>= 2000){
    return(1)
  }
return(0)
}
num_sim <- 1000
double_sim <- sapply(1:num_sim, play_double_or_nothing_roulette)
#Probability of doubling your money
sum(double_sim)/1000

```

```

<br />