### 2023-09-26 STAT405Project.qmd
9/26 finished

---
title: "Healthcare Insights and Sociodemographic Analysis of New York: A Comprehensive Analysis of Patient Records in 2013"
format: pdf
author: Abel Limachi, Austin Tran, Joel Villarino, and Krish Kumar (Group 3)
editor: visual
---
*(last updated 09/23/2023)*

**Introduction:**

In the context of the ever-evolving landscape of healthcare, our choice of dataset stems from the pressing need to delve into an in-depth analysis of the healthcare dynamics prevailing in today's society. New York, being one of the most densely populated states in the United States, presents a unique microcosm of diverse populations and a wide spectrum of urban and rural healthcare challenges. Motivated by the significance of understanding these complexities, our research group has selected a comprehensive dataset encompassing 2,428,905 patient records, representing healthcare interactions across various hospitals throughout New York in the year 2013.

This rich dataset spans a spectrum of information, including patient demographics, detailed medical records, financial aspects related to their hospitalization, and critical contextual information pertinent to each patient's stay. This extensive dataset holds immense potential for generating compelling visualizations and conducting statistical analyses, offering valuable insights into the intricacies of the New York healthcare system while also providing broader insights that can contribute to the understanding of healthcare systems more generally.

The following project contains different visualizations and statistics that will provide you more information on the different social and economic factors associated with patients in the state of New York in 2013.

**Dataset:**

2013 Hospital Inpatient Discharges in the State of New York

*The following data set was received from the official New York State Health website on heath.data.ny.gov*

<https://health.data.ny.gov/Health/Hospital-Inpatient-Discharges-SPARCS-De-Identified/npsr-cm47/data>

**From website:**

The Statewide Planning and Research Cooperative System (SPARCS) Inpatient De-identified File contains discharge level detail on patient characteristics, diagnoses, treatments, services, and charges. This data file contains basic record level detail for the discharge. The de-identified data file does not contain data that is protected health information (PHI) under HIPAA. The health information is not individually identifiable; all data elements considered identifiable have been redacted. For example, the direct identifiers regarding a date have the day and month portion of the date removed.

Discharges that are identified as Abortion will have some information redacted; Facility ID, Facility Name, Health Service Area, Hospital County, Operating Certificate, and all provider license numbers. Patient zip code is limited to the first three digits. Zip codes for population size less than 20,000 are blank. Out of State are "OOS'. Patient age is presented in age groups: 0 to 17, 18 to 29, 30 to 49, 50 to 69, and 70 or Older.

**Logistics:**

-   2,428,905 rows of patient data

-   34 columns

    -   Health Service Area
    -   Hospital County
    -   Operating Certificate Number
    -   Facility ID, Facility Name
    -   Age Group, Zip Code (3 digit)
    -   Gender
    -   Race
    -   Ethnicity
    -   Length of Stay
    -   Type of Admission
    -   Patient Disposition
    -   Discharge Year
    -   CCS Diagnosis Code
    -   CCS Diagnosis Description
    -   CCS Procedure Code
    -   CCS Procedure Description
    -   APR DRG Code
    -   APR DRG Description
    -   APR MDC Code
    -   APR MDC Description
    -   APR Severity of Illness Code
    -   APR Severity of Illness
    -   APR Risk of Mortality
    -   APR Medical Surgical Description
    -   Payment Typology
    -   Payment Typology
    -   Birth Weight,
    -   Abortion Edit Indicator
    -   Emergency Department
    -   Total Charges
    -   Total Costs

# Visualizations and Plots

## **Plot 1: Price vs Length of Stay and Linear Regression Statistical Test**

<br />

```{r}
#| echo: false
#install.packages("readr")
library(readr)

data_set <- read_csv("Hospital_Inpatient_Discharges__SPARCS_De-Identified___2013.csv",show_col_types = FALSE)

cleaned_data <- na.omit(data_set)

```

<br />

```{r}
#| echo: false
y <- cleaned_data$`Total Charges`
```

```{r}
#| echo: false
length_of_stay <- strtoi(cleaned_data$`Length of Stay`)
plot(length_of_stay, y,  xlab = "Length of Stay", main = "Price vs Length of Stay", ylab = "Price")

lmfs <- lm(y ~ length_of_stay)
summary(lmfs)
abline(lmfs, col = "RED")
```

### Linear Regression Statistical Test and Analysis

In our analysis of patient data for the state of New York in 2013, we conducted a linear regression to explore the impact of the length of stay on the associated price. The results revealed a statistically significant and strong positive relationship between these variables. The coefficient for 'length_of_stay' was estimated at 6140.73 with a standard error of 6.622 and a remarkably high t-value of 927.39. This suggests that as the length of stay increases, there is a substantial increase in the price, a relationship supported by a p-value close to zero. The R-squared value of 0.5567 indicates that approximately 55.67% of the price variation can be explained by the length of stay, demonstrating its relevance as a predictor. These findings provide important insights for healthcare management and policymaking, highlighting the need to consider and manage the cost implications of varying lengths of stay for New York state patients in 2013.

### Plot Desciption and Inferencing

In addition to the regression analysis, we visualized the relationship between the length of stay and the associated price using a scatter plot with a fitted regression line. The scatter plot provides a clear graphical representation of the positive correlation between these variables. As the length of stay increases, there is a discernible upward trend in price. The regression line, which is based on the linear regression model we performed, further accentuates this trend. It serves as a useful tool for illustrating the direction and strength of the relationship, making it easier to interpret the practical implications of our findings. This visualization reinforces the statistical significance we observed in our analysis and enhances our understanding of the relationship between patient length of stay and healthcare costs for patients in the state of New York in 2013.

## **Plot 2: Boxplot of Total Costs**

```{r}
#| echo: false
cleaned_data$'Total Costs'[cleaned_data$'Total Costs' > quantile(cleaned_data$'Total Costs', 0.95)] <- quantile(cleaned_data$'Total Costs', 0.95)

summary(cleaned_data$'Total Costs')

boxplot(cleaned_data$'Total Costs', 
        main="Box Plot of Total Costs (Outliers Truncated)",
        ylab="Total Costs",
        horizontal = TRUE)
```

### Plot Desciption and Inferencing

The following boxplot refers to our Total Cost data and is representative of our data that has been cleaned to remove outliers that are in our top 5% of data points, meaning that the outliers that are extremely high are cleaned out. In our analysis, we used a horizontal box plot to summarize the distribution of 'Total Costs' in our dataset. The median, located at approximately \$8,247.96, divides the data into two halves, with the interquartile range (IQR) spanning from Q1 at \$4,519.54 to Q3 at \$15,840.81. Outliers, representing data points significantly deviating from the central distribution, are visible. The total cost range extends from \$0.41 to \$45,292.36, showcasing the data's spread. The mean total cost is around \$12,360.20, but outliers can influence this value. The boxplot also is skewed to the right. This visualization provides key insights into our cost structure analysis.

From this we can infer that there are many outliers, however, in the discussion of costs they are heavily spread and that most patients are paying within our IQR but there are many different outliers that spread when we are discussing costs for healthcare.

## **Plot 3: Payment Type Distribution**

```{r, warning = FALSE, fig.height = 6, fig.width = 50}
#| echo: false
options(repr.plot.width = 100, repr.plot.height = 100)

cleaned_data$`Payment Typology 1` <- as.factor(cleaned_data$`Payment Typology 1`)
payments <- table(cleaned_data$`Payment Typology 1`)
pie(payments, main = "Payment Type Distribution", col=rainbow(length(payments)), labels = NA, )
legend(1,1,1, legend = levels(cleaned_data$`Payment Typology 1`), fill = rainbow(length(payments)), cex = 0.7)

```

### Plot Desciption and Inferencing

The following piechart identifies the payment type distribution for patients in New York in 2013. As we can observe, most patients are using Medicare as a way to pay for their medical bills followed by Medicaid. These government programs are important to New York residents and provides important data on the impact of healthcare programs. The exact percentagse are Medicaid at 59.7% and Medicaid at 26.7%, followed by Private Insurance at around 10%. This information reveals how important the impact of governmental insurance programs are to individuals in New York.

## **Plot 4: Trends in Total Charges and Length of Stay**

```{r warning = FALSE}
#| echo: false

plot(cleaned_data$`Length of Stay`, cleaned_data$`Total Charges`, type = "l", col = "blue", xlab = "Length of Stay", ylab = "Total Charges", main = "Trends in Total Charges and Length of Stay")

lines(cleaned_data$`Birth Weight`, cleaned_data$`Total Costs`, col = "red")

legend("topright", legend = c("Total Charges", "Total Costs"), col = c("blue", "red"), lty = 1)

grid()
```

### Plot Desciption and Inferencing

The plotted data and corresponding projections unmistakably depict a consistent total cost profile, irrespective of variations in both the length of stay and total charges. This apparent lack of correlation between the length of stay and associated costs prompts a deeper inquiry into the potential contributing factors. It would be important to explore the influence of fixed operational costs, negotiated contractual agreements, and astute cost management strategies that may explain this observed financial consistancy between the costs for a hospital and for its patients

## **Plot 5: Barplot of Age Group Distribution of Patients**

```{r}
#| echo: false
ageGroups <- table(cleaned_data$`Age Group`)
barplot(ageGroups, main = "Age Group Distribution", xlab = "Age Group", 
        ylab = "Frequency", col = "green", border = "black", ylim = c(0, 300000)) 
```

### Plot Desciption and Inferencing

Another important piece of data that we wanted to observe is the amount of patients from each age group of the population. The data is divided into five segmests 0-17, 18-29, 30-49, 50-69, and 70-older. As we can see there is an upward trend of the amount of patients as the ages of the patients increases, although there is a small decrease between 0-17 and 8-29. The largest frequency if the 70-older age bucket which has around 289248 compared to 18-29 with our smallest frequency of 53009.This increase in older patients reveals the impact of older populations in New York and the necesity to increase health care efforts towards older populations.

## **Plot 6: Heatmap of Gender vs. Age Group**

```{r}
#| echo: false
heatmap_data <- table(cleaned_data$`Age Group`, cleaned_data$Gender)

row_names <- as.factor(rownames(heatmap_data))
col_names <- as.factor(colnames(heatmap_data))

heatmap(heatmap_data,
        Rowv = NA,          
        Colv = NA,       
        col = colorRampPalette(c("white", "blue"))(50),  
        xlab = "Gender",    
        main = "Gender v. Age Group",  
        labRow = row_names, 
        labCol = col_names)

```

### Plot Desciption and Inferencing

The following heatmap shows the relationship between the gender and age group for patients in New York in 2013. As we can observe, women tend to be patients at hospitals at a higher rate then men. Especially for the 70 or older bracket which is the highest frequency of hospital patients, females are significantly a larger proprotion of who are patients compared to men. Another important data point to understand is the identification of "U" or individuals who do not identify as male and female, which is high for ages 18-29, which is an important data point to help increase healthcare for individuals with fluid gender identity or may identify as non-binary.

<br />

## Plot 7: Stacked Bar Chart by Gender and Race

```{r}
#| echo: false
#install.packages("ggplot2")
library(ggplot2)
ggplot(cleaned_data, aes(x = Gender, fill = Race)) +
  geom_bar() +
  labs(
    title = "Stacked Bar Chart by Gender and Race",
    x = "Gender",
    y = "Count"
  ) +
  scale_fill_manual(
    values = c("White" = "lightblue", "Black" = "lightcoral", "Other" = "lightgreen"),
    name = "Race"
  ) +
  scale_y_continuous(labels = scales::comma)
  theme_minimal()

```

## **Plot 8:
```{r}
#| echo: false
#install.packages("ggplot2")
library(ggplot2)
ggplot(cleaned_data) + 
        aes("Race", "Total Costs", color = "salmon") + 
        geom_jitter()

```