### 2023-11-28 REVEALJS.qmd
Add files via upload

---
title: "Healthcare Analysis in the State of New York - Hospital Data"
subtitle: "STAT 405 Final Project" 
author: Krish Kumar, Abel Limachi, Austin Tran, Joel Villarino
format:
  revealjs: 
    transition: slide
    background-transition: fade
    transition-speed: slow
    center: true
    # embed-resources: true
    slide-number: true
    chalkboard: 
      buttons: false
    preview-links: auto
    css: styles.css
    footer: STAT 405 - Fall 2023 (Kumar, Limachi, Tran, Villarino)
server: shiny
---

# Primary Dataset Overview {background="#D2B48C"}

```{r setup, echo=FALSE,  warning = FALSE, message = FALSE, fig.align='center'}
#| context: setup

library(stringr)
library(dplyr)
library(grid)
library(gridExtra)
library(RSQLite)
library(shiny)
library(DBI)
library(RSQLite)
library(ggplot2)
library(sf)
library(waffle)
library(patchwork)
library(tigris)
library(ggplot2)
library(viridis)
library(lubridate)
library(sf)
library(DBI)
library(flextable)

dcon <- dbConnect(SQLite(), dbname = "finalDB.db", timeout = 10)
```

## 2013 New York Hospital Data

**Number of Patients**: $2,428,905$

Patient Data from the State of New York were aggregated in this dataset from the NY Health Dept. with features including:

**Variables:**

::: columns
::: {.column width="50%"}
::: {style="font-size:16px"}
-   APR Severity of Illness Code
-   APR Severity of Illness
-   Health Service Area
-   Hospital County
-   Age Group
-   Gender
-   Race
-   Ethnicity
-   Length of Stay
-   Total Charges
-   Total Costs
-   Payment Typology
-   CCS Procedure Description
-   CCS Diagnosis Code
:::
:::

::: {.column width="50%"}
::: {style="font-size:16px"}
-   CCS Diagnosis Description
-   CCS Procedure Code
-   APR MDC Code
-   APR MDC Description
-   APR DRG Code
-   APR DRG Description
-   APR Risk of Mortality
-   Abortion Edit Indicator
-   Emergency Department
-   Type of Admission
-   Facility ID
-   Facility Name
-   Zip Code (3 digit)
:::
:::
:::

## Variable Analysis

::: {style="font-size:25px"}
-   APR Severity of Illness
-   Hospital County
-   Age Group
-   Gender
-   Race
-   Ethnicity
-   Length of Stay
-   Total Charges
-   Total Costs
-   Payment Typology
:::

# Severity of Illness Visualizations {background="#D2B48C"}

## Waffle Plot of Severity of Illness by Race

```{r Waffle Plot}
query <- dbSendQuery(dcon,"
SELECT Race, APRSeverityofIllnessCode FROM '2013data';
                     ")
IllnessCodes <- dbFetch(query, -1)
dbClearResult(query)

waffle_data1 <- data.frame(count = table(IllnessCodes$APRSeverityofIllnessCode[IllnessCodes$Race == "Black/African American"])/2480)
waffle_data2 <- data.frame(count = table(IllnessCodes$APRSeverityofIllnessCode[IllnessCodes$Race == "White"])/5050)

layout(matrix(c(1, 2), nrow = 1))

plot1 <- waffle(waffle_data1, rows = 8, size = .5, 
            colors = c("#FFEDA0", "#FEB24C", "#FC4E2A", "#BD0026"), title = "Black/African Americans")
plot2 <- waffle(waffle_data2, rows = 8, size = .5, 
            colors = c("#FFEDA0", "#FEB24C", "#FC4E2A", "#BD0026"), title = "White")

arrange.plots <- plot1 + theme(plot.title = element_text(size = 10)) + plot2 + theme(plot.title = element_text(size = 10))
arrange.plots
```

## Violin Plot of Severity of Illness by Geographical Location

```{r Violin Plot}
query <- dbSendQuery(dcon,"
SELECT *
FROM '2013data';
")
cleaned_data <- dbFetch(query, -1)
dbClearResult(query)
subsetted_data <- cleaned_data[cleaned_data$`HealthServiceArea` %in% c("New York City", "Western NY"), ]

library(ggplot2)

# Create a violin plot for APR Severity of Illness Code
ggplot(subsetted_data, aes(x = factor(`HealthServiceArea`), y = `APRSeverityofIllnessCode`, fill = `HealthServiceArea`)) +
  geom_violin() +
  labs(
    x = "Health Service Area",
    y = "APR Severity of Illness Code",
    fill = "Health Service Area",
    title = "Distribution of APR Severity of Illness Code in New York City and Western NY"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

# Disparities in Charges and Costs {background="#D2B48C"}

## Bar Charts of Total Charge for Pregnant Women for Race (2013 & 2021)

```{r, .center}
library(ggplot2)
library(patchwork)

pregnant_womenQ <- dbSendQuery(dcon, "
    SELECT Race, TotalCharges
    FROM '2013data'
    WHERE APRMDCCode == 14 OR APRMDCCode == 15;")
pregnant_womenDF = dbFetch(pregnant_womenQ, -1)
dbClearResult(pregnant_womenQ)

plot_race2013 <- ggplot(pregnant_womenDF) + 
  aes(x = Race, y = TotalCharges) +
  stat_summary(fun = "mean", geom = "bar", fill = "salmon") +  # Calculate and plot averages
  labs(x = "Race", y = "Average Total Charges", title = "Average Total Charge by Race in 2013") +  # Customize axis labels
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size = 7))  # Rotate x-axis labels for

pregnant_womenQ2 <- dbSendQuery(dcon, "
    SELECT Race, TotalCharges
    FROM '2021data'
    WHERE APRMDCCode == 14 OR APRMDCCode == 15 AND Race != 'Multi-racial';")
pregnant_womenDF2 = dbFetch(pregnant_womenQ2, -1)
dbClearResult(pregnant_womenQ2)

plot_race2021 <- ggplot(pregnant_womenDF2) + 
  aes(x = Race, y = TotalCharges) +
  stat_summary(fun = "mean", geom = "bar", fill = "violet") +  # Calculate and plot averages
  labs(x = "Race", y = "Average Total Charges", title = "Average Total Charge by Race in 2021") +  # Customize axis labels
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size = 7))  # Rotate x-axis labels for

combined_plot_race <- plot_race2013 + plot_race2021
combined_plot_race
```

## Facet Density Plots of Total Cost per Age-Group Buckets

```{r}

gg <- ggplot(cleaned_data, aes(x = `TotalCosts`, y = ..density.., fill = `AgeGroup`)) +
  geom_density(alpha = 0.5) +
  facet_grid(`AgeGroup` ~ .) +
  labs(title = "Density Plot of Total Costs per Age-Group Bucket", x = "Total Costs", y = "Density")

gg + 
  scale_x_continuous(limits = c(0, 35000)) +
  scale_y_continuous(labels = scales::number_format(scale = 1e-3)) +
  labs(title = "Density Plot of Total Costs per Age-Group Bucket", x = "Total Costs", y = "Density (in thousands)") +
  theme_minimal()
```

# Payment Visualizations {background="#D2B48C"}

## Pie Chart of Payment Type

```{r}
par(mar = c(5, 4, 4, 2) + 0.1)

cleaned_data$`PaymentTypology1` <- as.factor(cleaned_data$`PaymentTypology1`)
payments <- table(cleaned_data$`PaymentTypology1`)
pie(payments, main = "Payment Type Distribution", col = rainbow(length(payments)), labels = NA)
legend(1, 1, 1, legend = levels(cleaned_data$`PaymentTypology1`), fill = rainbow(length(payments)), cex = 0.7)

```

## Pie Charts of Payment Type by Race

```{r Pie Charts of Payment Type by Race}
library(ggplot2)
library(gridExtra)
library(patchwork)

paymentTypeAndRaceQ <- dbSendQuery(dcon, "
                                    SELECT Race, PaymentTypology1 FROM '2013data' WHERE Race != 'Unknown';")
paymentTypeAndRaceDF <- dbFetch(paymentTypeAndRaceQ, -1)
dbClearResult(paymentTypeAndRaceQ)
#str(paymentTypeAndRaceDF)

pie_charts <- lapply(unique(paymentTypeAndRaceDF$Race), function(race) {
  data_subset <- subset(paymentTypeAndRaceDF, Race == race)
  pie_chart <- ggplot(data_subset, aes(x = "", fill = PaymentTypology1)) +
    geom_bar(width = 1) +
    coord_polar("y") +
    ggtitle(race) +
    theme_void() +
    scale_fill_brewer(palette = "Set3") +
    theme(legend.position = ifelse(race == 'Black/African American', "left", "none"),
          axis.text = element_text(size = 0),
          legend.text = element_text(size = 5)) +
    labs(fill = "Payment Type")

  return(pie_chart)
})
pie_charts[[1]] + pie_charts[[2]] + pie_charts[[3]] + plot_layout(ncol = 1)
```

# Analyzing Medicaid Recipients by Race and Borough in New York City {background="#D2B48C"}

## Heat Map

```{r Heat Map}
query <- dbSendQuery(dcon, "
  SELECT `HospitalCounty`, Race, COUNT(*) as count
  FROM `2013data`
  WHERE `HealthServiceArea` = 'New York City' 
    AND Race IN ('Black/African American', 'White')
  GROUP BY `HospitalCounty`, `Race`
")

data <- dbFetch(query, -1)

ggplot(data, aes(x = `HospitalCounty`, y = `Race`, fill = count)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "red") +
  labs(title = "Number of White and Black/Other Race Individuals on Medicaid in NYC", 
       x = "Borough", 
       y = "Race", 
       fill = "Count") +
  theme_minimal()

dbClearResult(query)

# Query the data
query <- dbSendQuery(dcon,"
  SELECT `HospitalCounty`, Race, COUNT(*) as count
  FROM `2013data`
  WHERE `HealthServiceArea` = 'New York City' 
    AND Race IN ('Black/African American', 'White')
  GROUP BY `HospitalCounty`, `Race`
"
)
data <- dbFetch(query, -1)

dbClearResult(query)

# Retrieve the NYC boroughs shape data and filter for the five boroughs
options(tigris_use_cache = TRUE)

# Retrieve the NYC boroughs shape data and filter for the five boroughs
ny_counties <- tigris::counties(state = "NY", cb = TRUE) %>%
  st_as_sf()

nyc_boroughs <- ny_counties[ny_counties$NAME %in% c("Bronx", "Kings", "Manhattan", "Queens", "Richmond"), ]

# Merge the data with the nyc_boroughs spatial object
data_sf <- merge(nyc_boroughs, data, by.x = "NAME", by.y = "HospitalCounty")

# Compute the centroids for the NYC boroughs for plotting circles
data_sf$centroid <- st_centroid(st_geometry(data_sf))

```

## New York City Map

```{r New York City Map}

db_path <- "finalDB.db"
con <- dbConnect(SQLite(), db_path)

# Query the data
query <- "
  SELECT `HospitalCounty`, Race, COUNT(*) as count
  FROM `2013data`
  WHERE `HealthServiceArea` = 'New York City' 
    AND Race IN ('Black/African American', 'White')
  GROUP BY `HospitalCounty`, `Race`
"
data <- dbGetQuery(con, query)

# Replace 'Manhattan' with 'New York' to match the shapefile data
data$HospitalCounty <- gsub("Manhattan", "New York", data$HospitalCounty)

# Disconnect from the database
dbDisconnect(con)

# Retrieve the NYC boroughs shape data and filter for the five boroughs
options(tigris_use_cache = TRUE)
ny_counties <- tigris::counties(state = "NY", cb = TRUE) %>% st_as_sf()

nyc_boroughs <- ny_counties[ny_counties$NAME %in% c("Bronx", "Kings", "New York", "Queens", "Richmond"), ]

# Merge the data with the nyc_boroughs spatial object
data_sf <- merge(nyc_boroughs, data, by.x = "NAME", by.y = "HospitalCounty", all.x = TRUE)

# Compute the centroids for the NYC boroughs for plotting circles
data_sf$centroid <- st_centroid(data_sf)

# Create a column with the X and Y coordinates of the centroids
data_sf$X <- st_coordinates(data_sf$centroid)[, "X"]
data_sf$Y <- st_coordinates(data_sf$centroid)[, "Y"]

# Plotting side-by-side maps
ggplot(data_sf) +
  geom_sf(aes(fill = NAME), color = "grey60", size = 0.5) +
  geom_point(aes(x = X, y = Y, size = count, color = Race), alpha = 0.7, show.legend = "point") +
  facet_wrap(~Race, ncol = 2) +
  scale_color_manual(values = c("Black/African American" = "blue", "White" = "red"), name = "Race") +
  scale_size_continuous(name = "Count", breaks = c(10000,20000,30000,40000,50000)) +
  labs(title = "Number of Black vs. White Individuals on Medicaid in NYC Boroughs",
       subtitle = "Size of point indicates count, color indicates county") +
  theme_minimal() +
  theme(legend.position = "right", axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank())

```

# Human Body Economic Diagram (HBED) {style="font-size: 30px"}

```{r}
#| context: setup
library(shiny)
library(grid)
library(gridExtra)
library(scales)
library(dplyr)
library(RSQLite)
library(readr)
library(dplyr)
library(xtable)

dcon <- dbConnect(SQLite(), dbname = "finalDB.db", timeout = 10)
query <- dbSendQuery(dcon,"
  SELECT *
  FROM '2021data';
  ")
  data_2021 <- dbFetch(query, -1)
  dbClearResult(query)
  query <- dbSendQuery(dcon,"
  SELECT *
  FROM '2013data';
  ")
  data_2013 <- dbFetch(query, -1)
  dbClearResult(query)
  cleaned_data <- data_2013
```

```{r}
#| panel: sidebar
titlePanel("Interactive Human Figure")

  # Dropdown menu for selecting body parts
selectInput("bodyPart", "Select a Body Part:", choices = c("Total","Head", "Torso", "Chest"))
selectInput("time", "Select a Time Period:",
              choices = c("2013", "2021"))
```

```{r}
#| panel: fill
# Output plot created with grid graphics
plotOutput("gridPlot", width = "600px", height = "500px")
```

```{r}
#| context: server

  # This function wraps existing grid drawing code
  drawYourPlot <- function() {
    if(input$time == "2013"){
   cleaned_data <- data_2013
}
  if(input$time == "2021"){
  cleaned_data <- data_2021
}

  chest <- select(cleaned_data, APRMDCCode, TotalCharges, LengthofStay) %>% 
  filter(APRMDCCode %in% c(16, 5, 8, 4, 9, 21))

head <- select(cleaned_data, APRMDCCode, TotalCharges,LengthofStay) %>% 
  filter(APRMDCCode %in% c(20, 2, 1, 3, 10, 18, 19, 23))

stomach <- select(cleaned_data, APRMDCCode, TotalCharges, LengthofStay) %>% 
  filter(APRMDCCode %in% c(6, 13, 7, 11, 12, 17, 15, 24))

chest_count <- length(chest$APRMDCCode)
head_count <- length(head$APRMDCCode)
stomach_count <- length(stomach$APRMDCCode)
total_count <- length(cleaned_data$APRDRGCode)

draw_bed<- function(){
#bed frame
grid.draw(rectGrob(x = 0.80, y = 0.3, width = 0.375, height = 0.1, gp = gpar(fill = "brown3", col = "brown3", lwd = 2)))
#mattress
grid.draw(rectGrob(x = 0.80, y = 0.325, width = 0.35, height = 0.12, gp = gpar(fill = "gray", col = "gray", lwd = 2)))
#pillow
grid.draw(rectGrob(x = 0.675, y = 0.38, width = 0.125, height = 0.05, gp = gpar(fill = "gray", col = "darkgray", lwd = 2)))

#left headboard
grid.draw(rectGrob(x = 0.6, y = 0.35, width = 0.05, height = 0.35, gp = gpar(fill = "brown3", col = "brown3", lwd = 2)))

#right headboard
grid.draw(rectGrob(x = 0.975, y = 0.325, width = 0.05, height = 0.3, gp = gpar(fill = "brown3", col = "brown3", lwd = 2)))

#blanket
grid.draw(rectGrob(x = 0.85, y = 0.325, width = 0.2, height = 0.135, gp = gpar(fill = "lightpink", col = "pink2", lwd = 2)))

grid.text("Average Charge per Day", x = 0.8, y = 0.5, just = "centre")

if(input$bodyPart == "Head"){
  grid.text(round(sum(head$TotalCharges)/sum(head$LengthofStay),2), x = 0.85, y = 0.33, just = "centre")
}
else if(input$bodyPart == "Chest"){
   grid.text(round(sum(chest$TotalCharges)/sum(chest$LengthofStay), 2), x = 0.85, y = 0.33, just = "centre")
}
else if(input$bodyPart == "Torso"){
   grid.text(round(sum(stomach$TotalCharges)/sum(stomach$LengthofStay), 2), x = 0.85, y = 0.33, just = "centre")
  }
}
head_gp <- gpar(fill = "white")
body_gp <- gpar(fill = "black")
arms_gp <- gpar(fill = "white")
legs_gp <- gpar(fill = "khaki")
if(input$bodyPart == "Total"){
grid.newpage()
color_palette <- colorRampPalette(c("green", "red"))
#Drawing the gradient key
grad_key_height <- 0.5
grad_key_width <- 0.05
grad_key_x <- 0.7
grad_key_y <- 0.5

draw_vertical_rectangle <- function(x, y, width, height, gp = gpar()) {
  grid.rect(x = x - width / 2, y = y - height / 2, width = width, height = height, gp = gp)
}

grid.rect(x = grad_key_x-.025, y = grad_key_y-0.005, width = grad_key_width, height = grad_key_height, gp = gpar(col = "black"))

colors <- color_palette(100)
for (i in 1:100) {
  rect_gp <- gpar(fill = colors[i], col = NA)
  y_pos <- grad_key_y - (grad_key_height / 2) + ((i - 1) / 100) * grad_key_height
  draw_vertical_rectangle(x = grad_key_x, y = y_pos, width = grad_key_width, height = grad_key_height / 100, gp = rect_gp)
}
# Add labels to the gradient key
grid.text("$", x = grad_key_x-0.015, y = grad_key_y - (grad_key_height / 2) - 0.02, just = "centre")
grid.text("$$$", x = grad_key_x-0.015, y = grad_key_y + (grad_key_height / 2) + 0.02, just = "centre")
grid.text("Costs", x = grad_key_x-0.015, y = grad_key_y + (grad_key_height / 2) + 0.1, just = "centre")

get_color <- function(count, max_count) {
  # Assuming 'color_palette' is a function or vector of colors that returns a color based on the index
  normalized_count <- count/max_count
  colors <- colorRampPalette(c("green", "yellow", "orange", "orange", "red"))(100)  # Blue for less, Red for more
  color_index <- max(1, ceiling(normalized_count*100))
  return(colors[color_index])
}
draw_circle <- function(x, y, radius, gp = gpar()) {
  grid.circle(x = x, y = y, r = radius, gp = gp)
}

max_group_cost <- max(mean(head$TotalCharges), mean(chest$TotalCharges), mean(stomach$TotalCharges))
head_count_gp <- gpar(fill = get_color(mean(head$TotalCharges), max_group_cost))
chest_count_gp <- gpar(fill = get_color(mean(chest$TotalCharges), max_group_cost))
stomach_count_gp <- gpar(fill = get_color(mean(stomach$TotalCharges), max_group_cost))

draw_vertical_rectangle(x = 0.3, y = 0.75, width = 0.1, height = 0.35, gp = arms_gp) # this is neck
draw_circle(x = 0.25, y = 0.78, radius = 0.11, gp = head_gp) # this is head
draw_vertical_rectangle(x = 0.35, y = 0.65, width = 0.2, height = 0.35, gp = body_gp) # this is body
draw_vertical_rectangle(x = 0.15, y = 0.65, width = 0.05, height = 0.35, gp = arms_gp) # this is left arm

draw_circle(0.25,0.78, (head_count/total_count)/10, head_count_gp)
draw_circle(0.30,0.60, (chest_count/total_count)/10, chest_count_gp)
draw_circle(0.25,0.43, (stomach_count/total_count)/10, stomach_count_gp)

draw_circle(0.85,.28, .03)
draw_circle(0.85,.36, .04)
draw_circle(0.85,.45, .045)
draw_circle(0.85,.555, .0525)
draw_circle(0.85,.68, .06)
grid.text("Less", x = 0.85, y = grad_key_y - (grad_key_height / 2) - 0.02, just = "centre")
grid.text("More", x = 0.85, y = grad_key_y + (grad_key_height / 2) + 0.02, just = "centre")
grid.text("# of Cases", x = 0.85, y = grad_key_y + (grad_key_height / 2) + 0.1, just = "centre")

draw_vertical_rectangle(x = 0.4, y = 0.65, width = 0.05, height = 0.35, gp = arms_gp) # this is right arm
draw_vertical_rectangle(x = 0.25, y = 0.35, width = 0.1, height = 0.3, gp = legs_gp) # this is left leg
draw_vertical_rectangle(x = 0.35, y = 0.35, width = 0.1, height = 0.3, gp = legs_gp) # this is right leg
grid.ellipse(x = 0.25, y = 0.88, size = 9, ar = 4, angle = 0, gp = gpar(fill = "black"))
}
    # Here you would add additional grid drawing code based on the input$bodyPart
    # For example:
    else if (input$bodyPart == "Head") {
      # Add grid elements specific to the head
      grid.newpage()

grid.circle(x = 0.25, y = 0.5, r = 0.35, gp = gpar(fill = "white", col = "black", lwd = 3))
grid.lines(x = c(0.15), y = c(0,0.19), 
           gp = gpar(col = "black", lwd = 3))
grid.lines(x = c(0.35), y = c(0,0.19), 
           gp = gpar(col = "black", lwd = 3))

colors <- colorRampPalette(c("red", "yellow","green"))(1000)
num_circles <- length(colors)

# CIRCLE
circle_radius <- 0.1
x_center <- 0.80
y_center <- 0.75

# GRADIETN FOR CIRCLE
for (i in 1:num_circles) {
  color <- colors[i]
  angle <- seq(0, pi, length.out = num_circles + 1)[i]
  x <- x_center + circle_radius * cos(angle) 
  y <- y_center + circle_radius * sin(angle) 
  #grid.draw()
  grid.circle(x = x, y = y, r = 50 / num_circles, gp = gpar(fill = color, col = color))
}
  #grid.rect(x = 0.3, y = 0.25, gp = gpar(fill = "white", col = "white"), width = 2, height = .5)
  # LINES for APR Code
# 4
grid.lines(x = c(0.8, 0.9), y = c(0.7, .8), 
           gp = gpar(col = "black", lwd = 3))

grid.ellipse(x = 0.25, y = 0.85, size = 30, ar = 4, angle = 0, gp = gpar(fill = "black"))
grid.text("APR Severity Level", x= 0.8, y = 0.95, just = "centre")
draw_bed()

    } else if (input$bodyPart == "Torso") {
      # Add grid elements specific to the torso
      grid.newpage()

      draw_circle <- function(x, y, radius, gp = gpar()) {
  grid.circle(x = x, y = y, r = radius, gp = gp)
}
draw_vertical_rectangle <- function(x, y, width, height, gp = gpar()) {
  grid.rect(x = x - width / 2, y = y - height / 2, width = width, height = height, gp = gp)
}
draw_vertical_rectangle(x = 0.45, y = 0.95, width = 0.4, height = 0.6, gp = body_gp) # this is body
draw_vertical_rectangle(x = 0.25, y = 0.35, width = 0.2, height = 0.4, gp = legs_gp) # this is left leg
draw_vertical_rectangle(x = 0.45, y = 0.35, width = 0.2, height = 0.4, gp = legs_gp) # this is right leg

colors <- colorRampPalette(c("red", "yellow","green"))(1000)
num_circles <- length(colors)

# CIRCLE
circle_radius <- 0.1
x_center <- 0.80
y_center <- 0.75

# GRADIETN FOR CIRCLE
for (i in 1:num_circles) { 
  color <- colors[i]
  angle <- seq(0, pi, length.out = num_circles + 1)[i]
  x <- x_center + circle_radius * cos(angle) 
  y <- y_center + circle_radius * sin(angle) 
  grid.circle(x = x, y = y, r = 50 / num_circles, gp = gpar(fill = color, col = color))
}
# LINES for APR Code

  # 2
grid.lines(x = c(0.75, .8), y = c(0.85, .7), 
           gp = gpar(col = "black", lwd = 3))

grid.text("APR Severity Level", x= 0.8, y = 0.95, just = "centre")
draw_bed()
    } else if (input$bodyPart == "Chest") {
      # Add grid elements specific to the chest
      grid.newpage()
draw_circle <- function(x, y, radius, gp = gpar()) {
  grid.circle(x = x, y = y, r = radius, gp = gp)
}
draw_vertical_rectangle <- function(x, y, width, height, gp = gpar()) {
  grid.rect(x = x - width / 2, y = y - height / 2, width = width, height = height, gp = gp)
}
draw_vertical_rectangle(x = 0.4, y = 0.55, width = 0.3, height = 0.6, gp = body_gp) # this is body
draw_vertical_rectangle(x = 0.10, y = 0.55, width = 0.05, height = 0.4, gp = arms_gp) # this is left arm
draw_vertical_rectangle(x = 0.45, y = 0.55, width = 0.05, height = 0.4, gp = arms_gp) # this is right arm

colors <- colorRampPalette(c("red", "yellow","green"))(1000)
num_circles <- length(colors)

# CIRCLE
circle_radius <- 0.1
x_center <- 0.80
y_center <- 0.75

# GRADIENT FOR CIRCLE
for (i in 1:num_circles) { 
  color <- colors[i]
  angle <- seq(0, pi, length.out = num_circles + 1)[i]
  x <- x_center + circle_radius * cos(angle) 
  y <- y_center + circle_radius * sin(angle) 
  grid.circle(x = x, y = y, r = 50 / num_circles, gp = gpar(fill = color, col = color))
}
# LINES for APR Code
#3
grid.lines(x = c(0.8, .85), y = c(0.7, .85), 
           gp = gpar(col = "black", lwd = 3))

grid.text("APR Severity Level", x= 0.8, y = 0.95, just = "centre")
draw_bed()
    }
  }

  # Reactive plot output based on the selection
  output$gridPlot <- renderPlot({
    drawYourPlot()
  })

chest <- select(cleaned_data, APRMDCCode, TotalCharges) %>% 
  filter(APRMDCCode %in% c(16, 5, 8, 4, 9, 21))

head <- select(cleaned_data, APRMDCCode, TotalCharges) %>% 
  filter(APRMDCCode %in% c(20, 2, 1, 3, 10, 18, 19, 23))

stomach <- select(cleaned_data, APRMDCCode, TotalCharges) %>% 
  filter(APRMDCCode %in% c(6, 13, 7, 11, 12, 17, 15, 24))

chest_count <- length(chest$APRMDCCode)
head_count <- length(head$APRMDCCode)
stomach_count <- length(stomach$APRMDCCode)
total_count <- length(cleaned_data$APRDRGCode)
```

# Conclusion {style="font-size: 20"}

-   Disparities are examined on the basis of Race and Age
-   Between 2013 and 2021, costs and charges have significantly risen
-   Individuals who are Black/African American or from Other Races may not be receiving equitable healthcare
-   Large Amounts of Minorities Dependent on Government Programs

# Resources

2013 Dataset:

[https://www.health.ny.gov/statistics/vital_statistics/2013/](https://www.health.ny.gov/statistics/vital_statistics/2013/https://www.health.ny.gov/statistics/vital_statistics/2021/)

2021 Dataset:

[https://www.health.ny.gov/statistics/vital_statistics/2021/](https://www.health.ny.gov/statistics/vital_statistics/2013/https://www.health.ny.gov/statistics/vital_statistics/2021/)

# Questions {background="#D2B48C"}